A studyforrest extension, an annotation of words and phonemes in the German dubbed movie ``Forrest Gump'' and its audio-description

### 2020-04-17 paper/main.tex
overall revisions

% provide full affiliation information (including full institutional address, ZIP code and e-mail address) for all authors, and identify who is/are the corresponding author(s).
% EMAIL hier noch "kanonisch" unterbringen
% Madeburg habe ich nicht mit reingenommen, obwohl MD/Sachsen-Anhalt unten wg. Stipendium unter Funding kommt.
\author[1, 2]{Christian~O.~Häusler}
\author[1, 2]{Michael Hanke}

\affil[1]{Institute of Neuroscience and Medicine, Brain \& Behaviour (INM-7), Research Centre Jülich, Jülich, Germany}
\affil[2]{Institute of Systems Neuroscience, Medical Faculty, Heinrich Heine University, Düsseldorf, Germany}

\maketitle
\thispagestyle{fancy}
\begin{abstract}
% up to 300 words
Kommt ganz am Ende;
Intro, Method, Results, Discussion;
hier ggf. bereits Praat erwähnen;
im Gegensatz zu Sätzen und Phones haben Einzelworte zusätzliche Annos

\todo[inline]{The abstract should come now. It is needed to evaluate the
concept of the paper in the light of its content and vice versa.}

\end{abstract}
\section*{Keywords}
% maximal 8
language, speech, annotation, fMRI, natural stimulation, narrative, studyforrest

@@

\section*{Introduction}
% Links sind jetzt im laufenden Text; habs nach wie vor mit \href gemacht
% sieht bei längeren Links natürlich kacke aus

% largely adopted from haeusler2016annotation
% https://f1000research.com/articles/5-2273
% In Einleitung muss wohl noch etwas zu "reproducible research" und "datalad dataset" kommen
Cognitive neuroimaging research is moving towards studying brain behavior
under conditions of real-life(-like complexity). Motion pictures \citep{hasson2008neurocinematics} and continuous narratives \citep{honey2012not, lerner2011topographic} are increasingly utilized as so called naturalistic stimuli.

% new; not from haeusler2016annotation
Because naturalistic stimuli are multi-dimensional and their temporal structure
is (usually) unknown, data-driven methods like intersubject correlation
(ISC)\citep{hasson2004intersubject} or independent component analysis
(ICA)\citep{bartels2004chronoarchitecture} are often used to analyze
corresponding fMRI data. Nevertheless, theory-driven\todo{model-based?, there is little theroy in GLM} analyses like a general
linear model (GLM) are necessary to not just explain how but also when and why a
brain region is responding to a stimulus \citep{hamilton2018revolution}. Studies
using GLMs based on annotations of a stimulus' temporal structure have
elucidated how the brain responds to visual features of a movie
\citep{bartels2004mapping} or speech-related features of a narrative
\citep{rocca2019language}.

% adapted from haeusler2016annotation
% nun ja, folgend nicht super korrekt: sentences sind nicht vollständige
% Sätze im linguistischen Sinne (Subjekt, Objekt, Prädikat), sondern auch
% Ausrufe ("Forrest!") oder mitunter isolierte non-speech "Sätze" ("oouh").
% Sätze im linguistischen Sinne sind in der Annotation durch "." oder "?" am
% Ende gekennzeichnet. Sollte aber so passen.
For this publication, we annotated the onsets and offsets, and content of
sentences, single words, and phonemes spoken in the movie Forrest Gump (R.
Zemeckis, Paramount Pictures, 1994) \citep{ForrestGumpMovie}. fMRI data of
participants watching the audio-visual movie \citep{hanke2016simultaneous} and
it's audio-description for visually impaired persons \citep{hanke2014audiomovie}
are the core data of the publicly available \textit{studyforrest} dataset
(\href{www.studyforrest.org}{studyforrest.org}). Additional fMRI data
\citep{sengupta2016extension} include retinotopic mappings and localization of
higher visual areas of the same participants.
% paper regarding mention musical not mentioned
% und damit zitiere ich mich als dortiger Coauthor nicht selbst. Skandal!

The current annotation facilitates modelling of hemodynamic brain responses
correlating with speech-related events in the movie and audio-description.
Further, it extends already available annotations of portrayed emotions
\citep{labs2015portrayed} and perceived emtions
\citep{lettieri2019emotionotopy}, as well as cuts and locations depicted in the movie \citep{haeusler2016annotation}. All annotations can be used in any study focusing on other aspects of real-life cognition by serving as additional
confound measures describing key properties of major building blocks of the stimulus.

\section*{Materials and methods}
\subsection*{Stimulus}
% References sind vom Anno-Paper in die .bib-Datei übernommen worden; die
% “hardcoded” Zitationen sind aber noch im laufenden Text.
We annotated speech in stimuli built from the motion picture ``Forrest Gump'' as
used in two previous studies: an audio-visual movie
\citep{hanke2016simultaneous}, and its temporally aligned audio-description
\citep{hanke2014audiomovie}. For the movie stimulus, we used the audio track of
the German DVD release (Paramount Home Entertainment, Germany, 2002; PAL video,
DE103519SV)\citep{ForrestGumpDVD}. For the audio-only stimulus, we used the
movie's audio-description that was broadcast as an additional audio track for
visually impaired listeners on Swiss public television (Koop, Michalski,
Beckmann, Meinhardt \& Benecke, produced by Bayrischer Rundfunk,
2009)\citep{ForrestGumpGermanAD}. The plot of the movie is already carried by an
off-screen voice of the main character Forrest Gump. In the largely identical audio-description, a male narrator additionally describes essential aspects of the visual scenery when there is no off-screen voice, dialog, or other relevant auditory content
\subsection*{Annotation procedure}
% wo kommen die "preliminaries transcriptions" her?
% Die "most basic" Datei, die das Dataset enthält ist der Merge aus Diaolog, Narrator & Non-Speech-Annotation
Preliminary, orthographic transcriptions of dialogues (german\_dialog.csv),
non-speech vocalizations (e.g. laughing or groaning;
non\_speech\_vocalization.json), and the audio-description's narrator
(german\_audio\_description.csv) were merged and converted to Praat's
\citep{boersma2019praat} .TextGrid format (
\href{www.fon.hum.uva.nl/praat}{fon.hum.uva.nl/praat}). The merged transcription
containing one onset and offset for usually a couple of sentences was opened
along (the audio file of) the audio-description in Praat v6.0.22 for manual
improvements: In several passes, the approximate temporal onsets and offsets
were corrected. Intervals containing several sentences were split into intervals
containing only one sentence. When two or more persons were speaking
simultaneously the less dominant (background) voice was dropped. Low volume
background speech (especially occurring during music or continuous environmental
noise) or low volume non-speech vocalizations that are difficult to recognize by
a regular audience were dropped, too.

% Alignment preparations
We used the Montreal Forced Aligner v1.0.1
\href{montrealcorpustools.github.io/Montreal-Forced-Aligner}{montrealcorpustools.github.io/Montreal-Forced-Aligner}\citep{mcauliffe2017montreal};to
algorithmically identify the onset and offset of each word and phoneme embedded
in the transcribed sentences. To enable the aligner to look up the pronunciation
of every word, we chose a German pronunciation dictionary
\href{https://raw.githubusercontent.com/prosodylab/prosodylab.dictionaries/master/de.dict}{raw.githubusercontent.com/prosodylab/prosodylab.dictionaries/master/de.dict}
provided by Prosodylab (\href{http://prosodylab.org}{prosodylab.org}) that uses
the Prosodylab PhoneSet. The dictionary was manually updated with movie-specific
German words. The pronunciation of English words occurring in the otherwise
German audio track was taken from an English pronunciation dictionary
(\href{http://mlmlab.org/mfa/dictionaries/english.dict}{mlmlab.org/mfa/dictionaries/english.dict} that uses the ARPAbet PhoneSet.

% Alignment
% Nicht erwähnt: Wenn die Korrektur der Phoneme unmöglich war, habe ich einfach
% die Event-Borders zwischen den Phonemen herausgenommen. Wortbeginn und Wortende stimmen dann mit einem neuen, "künstlich-großen Phoneme" überein.
% Das "Groß-Phonem" enthält dann ohne Grenzen zwischen den tatsächlichen Phonemen des Wortes alle Phoneme des Wortes ("es ist kompliziert zu erklären, aber im Grunde ganz einfach).
% kommt selten vor, ist im Grunde "noise" und ist in der Anno sofort als solches zu erkennen.
The audio file was converted from .FLAC to .WAV via ffmpeg v4.1.4
(\href{www.ffmpeg.org}{ffmpeg.org}) to meet the aligner's requirements. This
.WAV file, the transcription of sentences, and the updated dictionary were
submitted to the aligner that first trained an acoustic model on the data and
then performed the alignment. In a first step, the resulting .TextGrid file that
contains onsets and offsets of words and phonemes was opened in Praat to check
and correct the automatic alignment. In several passes, words [and their
corresponding phonemes] on which the automatic alignment performed moderately
were corrected. Some (low volume) sentences that are spoken in continuously
noisy settings (pursuit, battle, hurricane, demonstration/rally, disco) were
removed due to poor overall alignment performance. In a second step, the
complete sentences of the orthographic transcription were copied into the
annotation (s. table \ref{tab:overview} for an overview of the annotation's content). In a third step, the name of the speaker was added for every sentence (s. table \ref{tab:speakers} for the most often occurring speakers).
During every new step results from all previous steps were checked again for errors and corrected.

\begin{table*}[tbp]
\caption{Overview of the speech annotation's content for the whole stimulus and it's individual segments. The label ``Sentences''' comprises complete grammatical sentences (i.e. having a subject and a predicate, and additionally marked with a full stop in the annotation). The  ``Sentences'' also comprises exlamations or non-speech vocalizations in quick succession (e.g. ``ha, ha, ha'') or in isolation (e.g. ``ha'', at time points when speakers rapidly switch). The label ``Words'' comprises every single word or non-speech vocalization (N=\aPosNonspeechAll) in isolation.}
\label{tab:overview}
\begin{tabular}{llllllllll}
\toprule
\textbf{label} & \textbf{all} & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8}\\
\midrule
sentences  & \aSentencesAll & \aSentencesI & \aSentencesII & \aSentencesIII & \aSentencesIV & \aSentencesV & \aSentencesVI & \aSentencesVII & \aSentencesVIII \tabularnewline
words  & \aWordsAll & \aWordsI & \aWordsII & \aWordsIII & \aWordsIV & \aWordsV & \aWordsVI & \aWordsVII & \aWordsVIII \tabularnewline
phonemes  & \aPhonesAll & \aPhonesI & \aPhonesII & \aPhonesIII & \aPhonesIV & \aPhonesV & \aPhonesVI & \aPhonesVII & \aPhonesVIII \tabularnewline
\bottomrule
\end{tabular}
\end{table*}

% bei ein paar Namen bin ich ggf. von Labs abgewichen (genauer identifiziert, ggf. kleine Fehler korrigiert. Ist dann aber nur bei Personen, die 1-2x mit 1-2 Sätzen auftauchen.
\begin{table*}[btp]
\caption{Sentences spoken by the ten most often occurring speakers sorted alphabetically. The narrator only occurrs in the audio-description of the movie. Shown names are identical to the names used in \citep{labs2015portrayed}. Overall 97 persons were identified.}
\label{tab:speakers}
\begin{tabular}{llllllllll}
\toprule
\textbf{name} & \textbf{all} & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8}\\
\midrule
Bubba  & \aBubbaAll & \aBubbaI & \aBubbaII & \aBubbaIII & \aBubbaIV & \aBubbaV & \aBubbaVI & \aBubbaVII & \aBubbaVIII \tabularnewline
Forrest  & \aForrestAll & \aForrestI & \aForrestII & \aForrestIII & \aForrestIV & \aForrestV & \aForrestVI & \aForrestVII & \aForrestVIII \tabularnewline
Forrest (child)  & \aForrestchildAll & \aForrestchildI & \aForrestchildII & \aForrestchildIII & \aForrestchildIV & \aForrestchildV & \aForrestchildVI & \aForrestchildVII & \aForrestchildVIII \tabularnewline
Forrest (v.o.)  & \aForrestvoAll & \aForrestvoI & \aForrestvoII & \aForrestvoIII & \aForrestvoIV & \aForrestvoV & \aForrestvoVI & \aForrestvoVII & \aForrestvoVIII \tabularnewline
Hancock  & \aHancockAll & \aHancockI & \aHancockII & \aHancockIII & \aHancockIV & \aHancockV & \aHancockVI & \aHancockVII & \aHancockVIII \tabularnewline
Jenny  & \aJennyAll & \aJennyI & \aJennyII & \aJennyIII & \aJennyIV & \aJennyV & \aJennyVI & \aJennyVII & \aJennyVIII \tabularnewline
Jenny (child)  & \aJennychildAll & \aJennychildI & \aJennychildII & \aJennychildIII & \aJennychildIV & \aJennychildV & \aJennychildVI & \aJennychildVII & \aJennychildVIII \tabularnewline
Lt. Dan  & \aLtdanAll & \aLtdanI & \aLtdanII & \aLtdanIII & \aLtdanIV & \aLtdanV & \aLtdanVI & \aLtdanVII & \aLtdanVIII \tabularnewline
Mrs. Gump  & \aMrsgumpAll & \aMrsgumpI & \aMrsgumpII & \aMrsgumpIII & \aMrsgumpIV & \aMrsgumpV & \aMrsgumpVI & \aMrsgumpVII & \aMrsgumpVIII \tabularnewline
Narrator  & \aNarratorAll & \aNarratorI & \aNarratorII & \aNarratorIII & \aNarratorIV & \aNarratorV & \aNarratorVI & \aNarratorVII & \aNarratorVIII \tabularnewline
\bottomrule
\end{tabular}
\end{table*}

% descriptive Nouns
% hier ggf. noch angeben, dass die effektiv identisch waren
% Florian hat zwei Regeln falsch verstanden, sonst war übereinstimmung >90%
A further annotation of the audio-description's narrator was done manually: Two
persons performed a categorization of nouns embedded in sentences spoken by the
narrator to describe the movie's missing visual content. Categories focus on the
cinematographic scene's environment, inherent objects, persons, and a person's
appearance. A preliminary annotation according to the rules (s. ist jetzt in der obigen Tabelle aus Platzgründen nicht mehr drin; in dieser,\ref{tab:descr-nouns}, einer neuen Tabelle oder im laufenden Text [wird viel Text] bringen) was done by one person and corrected in several runs by the author.\todo{I thought we said that this aspect should move to the PPA paper?}
    \caption{Narrator's descriptive nouns. All categories sorted alphapetically. Examples are given in English. Description (of rules) fehlt in dieser Tabelle (vgl. alte Tabelle auf der allerletzten Seite). Wird nämlich in einer Tabelle mit horizontaler Ausrichtung eng. Kategorie ``++'' enthält ausnahmesweise auch adverbiale Best. der Zeit.}
\label{tab:descr-nouns}
\begin{tabular}{lllllllllll}
\textbf{label} & \textbf{examples} & \textbf{all} & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8} \\
body & back, hip, shoulder; jacket, dress, shirt & \aDescrBodyAll & \aDescrBodyI & \aDescrBodyII & \aDescrBodyIII & \aDescrBodyIV & \aDescrBodyV & \aDescrBodyVI & \aDescrBodyVII & \aDescrBodyVIII \tabularnewline
bodypart  & arm, finger, leg, toe & \aDescrBodypartAll & \aDescrBodypartI & \aDescrBodypartII & \aDescrBodypartIII & \aDescrBodypartIV & \aDescrBodypartV & \aDescrBodypartVI & \aDescrBodypartVII & \aDescrBodypartVIII\tabularnewline
face & face, ear, nose, mouth & \aDescrFaceAll & \aDescrFaceI & \aDescrFaceII & \aDescrFaceIII & \aDescrFaceIV & \aDescrFaceV & \aDescrFaceVI & \aDescrFaceVII & \aDescrFaceVIII \tabularnewline
female & nurse, mother, woman & \aDescrFemaleAll & \aDescrFemaleI & \aDescrFemaleII & \aDescrFemaleIII & \aDescrFemaleIV & \aDescrFemaleV & \aDescrFemaleVI & \aDescrFemaleVII & \aDescrFemaleVIII \tabularnewline
females & women & \aDescrFemalesAll & \aDescrFemalesI & \aDescrFemalesII & \aDescrFemalesIII & \aDescrFemalesIV & \aDescrFemalesV & \aDescrFemalesVI & \aDescrFemalesVII & \aDescrFemalesVIII \tabularnewline
fname & Jenny & \aDescrFnameAll & \aDescrFnameI & \aDescrFnameII & \aDescrFnameIII & \aDescrFnameIV & \aDescrFnameV & \aDescrFnameVI & \aDescrFnameVII & \aDescrFnameVIII \tabularnewline
furniture & bench, bed, table, chair & \aDescrFurnitureAll & \aDescrFurnitureI & \aDescrFurnitureII & \aDescrFurnitureIII & \aDescrFurnitureIV & \aDescrFurnitureV & \aDescrFurnitureVI & \aDescrFurnitureVII & \aDescrFurnitureVIII \tabularnewline
geo & building, tree, street, alley, meadow, cornfield, river & \aDescrGeoAll & \aDescrGeoI & \aDescrGeoII & \aDescrGeoIII & \aDescrGeoIV & \aDescrGeoV & \aDescrGeoVI & \aDescrGeoVII & \aDescrGeoVIII \tabularnewline
geo-room & living room; wall, door, window, floor, turf & \aDescrGeoroomAll & \aDescrGeoroomI & \aDescrGeoroomII & \aDescrGeoroomIII & \aDescrGeoroomIV & \aDescrGeoroomV & \aDescrGeoroomVI & \aDescrGeoroomVII & \aDescrGeoroomVIII \tabularnewline
head & head, hair, ear, neck, helmet & \aDescrHeadAll & \aDescrHeadI & \aDescrHeadII & \aDescrHeadIII & \aDescrHeadIV & \aDescrHeadV & \aDescrHeadVI & \aDescrHeadVII & \aDescrHeadVIII \tabularnewline
male & man, father, soldier & \aDescrMaleAll & \aDescrMaleI & \aDescrMaleII & \aDescrMaleIII & \aDescrMaleIV & \aDescrMaleV & \aDescrMaleVI & \aDescrMaleVII & \aDescrMaleVIII \tabularnewline
males & boys, opponents & \aDescrMalesAll & \aDescrMalesI & \aDescrMalesII & \aDescrMalesIII & \aDescrMalesIV & \aDescrMalesV & \aDescrMalesVI & \aDescrMalesVII & \aDescrMalesVIII \tabularnewline
mname & Bubba, Kennedy & \aDescrMnameAll & \aDescrMnameI & \aDescrMnameII & \aDescrMnameIII & \aDescrMnameIV & \aDescrMnameV & \aDescrMnameVI & \aDescrMnameVII & \aDescrMnameVIII \tabularnewline
object & telephone, car & \aDescrObjectAll & \aDescrObjectI & \aDescrObjectII & \aDescrObjectIII & \aDescrObjectIV & \aDescrObjectV & \aDescrObjectVI & \aDescrObjectVII & \aDescrObjectVIII \tabularnewline
objects & wheels, photos & \aDescrObjectsAll & \aDescrObjectsI & \aDescrObjectsII & \aDescrObjectsIII & \aDescrObjectsIV & \aDescrObjectsV & \aDescrObjectsVI & \aDescrObjectsVII & \aDescrObjectsVIII \tabularnewline
persons & hippies, patients & \aDescrPersonsAll & \aDescrPersonsI & \aDescrPersonsII & \aDescrPersonsIII & \aDescrPersonsIV & \aDescrPersonsV & \aDescrPersonsVI & \aDescrPersonsVII & \aDescrPersonsVIII \tabularnewline
setting\_new & on a ``bridge'', on an ``alley'', on ``campus'' & \aDescrSettingnewAll & \aDescrSettingnewI & \aDescrSettingnewII & \aDescrSettingnewIII & \aDescrSettingnewIV & \aDescrSettingnewV & \aDescrSettingnewVI & \aDescrSettingnewVII & \aDescrSettingnewVIII \tabularnewline
setting\_rec & at the ``bus stop'' & \aDescrSettingrecAll & \aDescrSettingrecI & \aDescrSettingrecII & \aDescrSettingrecIII & \aDescrSettingrecIV & \aDescrSettingrecV & \aDescrSettingrecVI & \aDescrSettingrecVII & \aDescrSettingrecVIII \tabularnewline
++ &  in the ``evening'', it's ``daytime'', ``later'' & \aDescrAll & \aDescrI & \aDescrII & \aDescrIII & \aDescrIV & \aDescrV & \aDescrVI & \aDescrVII & \aDescrVIII \tabularnewline
\todo[inline]{If the descriptive nounes are moved to the PPA paper, this table should be removed}

@@

@@

pis & substituting indefinite pronoun & \rPisAll & \rPisI & \rPisII & \rPisIII & \rPisIV & \rPisV & \rPisVI & pper \rPisVII & \rPisVIII \tabularnewline

@@

\todo[inline]{I think this table doesn't talk about 'regressors'. It talks about the
events that were selected/available to build GLM regressors from them. Hence the term
'event' should be used here, and cross-reference to figure \ref{fig:reg} should be made}
\todo[inline]{what does e1-12 stand for? It makes the label column too wide}

>>>>

\author[1, 2]{Christian~O.~Häusler \\ email \href{mailto:c.haeusler@fz-juelich.de}{c.haeusler@fz-juelich.de} }
\author[1, 2]{Michael~Hanke}

\affil[1]{Institute of Neuroscience and Medicine, Brain \& Behaviour (INM-7), Research Centre Jülich, Jülich, 52425, Germany}
\affil[2]{Institute of Systems Neuroscience, Medical Faculty, Heinrich Heine University, Düsseldorf, 40225, Germany}

\maketitle

\todo[inline]{wie email und ORCID (du: https://orcid.org/0000-0001-6398-6370; ich: https://orcid.org/0000-0002-0936-317X) unterbringen?}

\thispagestyle{fancy}
\begin{abstract}
% up to 300 words
\todo[inline]{The abstract should come now}
Here we present an annotation of speech in the audio-visual movie ``Forrest Gump'' and it's audio-only variant (i.e. audio-description), as an addition to a large public functional brain imaging dataset (\url{http://studyforrest.org}). The annotation provides information about the exact timing of each of the \aSentencesAll\ spoken sentences, \aWordsAll\ words (including \aPosNonspeechAll\ non-speech vocalizations), and \aPhonesAll\ phonemes.

Zusätzliche Annos der Worte

This annotation enables further studies of language perception and semantic processing under conditions of real-life complexity using the studyforrest dataset.

Wort zur Validierungsstudie?
\end{abstract}

\section*{Keywords}
% maximal 8
annnotation, language, speech, lateralization, narrative, natural stimulation, fMRI, studyforrest

@@

\section*{Introduction}
\todo[inline]{Bitte am Ende (?) der Einleitung etwas zu "reproducible research" \& "datalad dataset" bringen}
Cognitive neuroimaging research is moving towards studying brain behavior
under conditions of real-life-like complexity. Motion pictures \citep{hasson2008neurocinematics} and continuous narratives \citep{honey2012not, lerner2011topographic} are increasingly utilized as so called naturalistic stimuli.

Because naturalistic stimuli are multi-dimensional and their temporal structure
is usually unknown, data-driven methods like intersubject correlation (ISC)\citep{hasson2004intersubject} or independent component analysis
(ICA)\citep{bartels2004chronoarchitecture} are often used to analyze
corresponding fMRI data.
Nevertheless, model-driven methods like a general linear model (GLM) are necessary to not just explain how but also when (and why) a brain region is responding to a stimulus \citep{hamilton2018revolution}.
Studies using GLMs based on annotations of a stimulus' temporal structure have
elucidated how the brain responds to visual features of a movie
\citep{bartels2004mapping} or speech-related features of a narrative
\citep{rocca2019language}.

% hier sind “sentences” nicht vollständige Sätze im linguistischen Sinne (Subjekt, Objekt, Prädikat), sondern auch Ausrufe ("Forrest!") oder mitunter isolierte non-speech "Sätze" ("oouh").
% Sätze im linguistischen Sinne sind in der Annotation durch "." oder "?" am
% Ende gekennzeichnet.
For this publication, we annotated the onsets and offsets, and content of sentences, single words and phonemes spoken in the audio-visual (AV) movie ``Forrest Gump'' \citep{ForrestGumpMovie} and its corresponding audio-only (AO) variant (i.e. audio-description) for visually impaired persons \citep{ForrestGumpDVD}.
% paper regarding mention musical not mentioned
fMRI data of participants watching the audio-visual movie \citep{hanke2016simultaneous} and listening to the audio-description \citep{hanke2014audiomovie} are the core data of the publicly available \textit{studyforrest} dataset (\href{www.studyforrest.org}{studyforrest.org}).
Additional fMRI data include retinotopic mappings and localization of higher visual areas of the same participants \citep{sengupta2016extension}.

\todo[inline]{this work seems to improve/replace previous work (in part), it is important to state that clearly and identify which pieces are involved; welche previous work wird verbessert oder sogar replaced?}
The current annotation facilitates modelling of hemodynamic brain responses
correlating with speech-related events in the two stimuli. Further, it extends already available annotations of portrayed emotions \citep{labs2015portrayed}, perceived emotions \citep{lettieri2019emotionotopy}, as well as cuts and locations depicted in the movie \citep{haeusler2016annotation}. All annotations can be used in any study focusing on other aspects of real-life cognition by serving as additional confound measures describing key properties of major building blocks of the stimulus.

\section*{Materials and methods}
\subsection*{Stimulus}
We annotated speech in the slightly shortened ``research cut'' of the movie \citep{hanke2016simultaneous} and its temporally aligned audio-description \citep{hanke2014audiomovie}.
For the movie stimulus, we used the audio track of the German DVD release \citep{ForrestGumpDVD}.
For the audio-only stimulus, we used the movie's audio-description that was broadcast as an additional audio track for visually impaired listeners on Swiss public television \citep{ForrestGumpGermanAD}.
The plot of the movie is already carried by an off-screen voice of the main character Forrest Gump.
In the largely identical audio-description, a male narrator additionally describes essential aspects of the visual scenery when there is no off-screen voice, dialog, or other relevant auditory content.

\subsection*{Annotation procedure}
\todo[inline]{wo kommen die "preliminaries transcriptions" etc.? Habe die von Medusa; die "most basic" Datei, die das Dataset enthält ist wie abgesprochen der Merge aus Dialog (ungetrennte Sentences), Narrator \& Non-Speech-Annotation = \texttt{speech-vocal.csv}}
Preliminary, orthographic transcriptions of dialogues (\texttt{german\_dialog.csv}), non-speech vocalizations (e.g. laughing or groaning; \texttt{non\_speech\_vocalization.json}), and the audio-description's narrator
(\texttt{german\_audio\_description.csv}) were merged (\texttt{speech-vocal.csv}) and converted to Praat's \citep{boersma2019praat} TextGrid format  (\texttt{speech-vocal.TextGrid}).
% \href{www.fon.hum.uva.nl/praat}{fon.hum.uva.nl/praat}.
The merged transcription containing one onset and offset for usually a couple of sentences was edited in Praat v6.0.22 for manual validation against the original audio.
The following steps were done in several passes to iteratively improve the quality of the data:
Approximate temporal onsets and offsets were corrected, and intervals containing several sentences were split into intervals containing only one sentence.
\todo[inline]{what is a regular audience? without neurological disorders ;-); Simpel: eine audience, die den Film nicht 45x auf voller Lautstärke über Kopfhörer mit Transkription danebenliegend gehört hat ;-)}
When two or more persons were speaking simultaneously the less dominant voice was dropped.
Low volume background speech (especially occurring during music or continuous environmental noise) or low volume non-speech vocalizations that are difficult to recognize by the audience were dropped, too.

% Alignment preparations
We used the Montreal Forced Aligner v1.0.1 \citep{mcauliffe2017montreal} to
algorithmically identify the onset and offset of each word and phoneme embedded
in the transcribed sentences.
% http://prosodylab.org
% https://raw.githubusercontent.com/prosodylab/prosodylab.dictionaries/master/de.dict
To enable the aligner to look up the phonemes within every word, we chose a German pronunciation dictionary provided by Prosodylab \citep {gorman2011prosodylab} that uses the Prosodylab PhoneSet to describe the pronounciation of single phonemes. To improve the quality of the automatic alignment, this dictionary was manually updated with German words that were missing in the dictionary but occur in the stimuli. The pronunciation of English words occurring in the otherwise German audio track was taken from an English pronunciation dictionary provided by the Montreal Language Modeling lab (\href{http://mlmlab.org/mfa/dictionaries/english.dict}{mlmlab.org/mfa/dictionaries/english.dict} that uses the ARPAbet PhoneSet.

% Alignment
\todo[inline]{Nicht erwähnt: Wenn die Korrektur der Phoneme unmöglich war, habe ich einfach die Event-Borders zwischen den Phonemen herausgenommen. Wortbeginn und Wortende stimmen dann mit einem neuen, "künstlich-großen Phoneme" überein.
Das "Groß-Phonem" enthält dann ohne Grenzen zwischen den tatsächlichen Phonemen des Wortes alle Phoneme des Wortes ("es ist kompliziert zu erklären, aber im Grunde ganz einfach). kommt jedoch so selten vor, ist im Grunde "noise" und ist in der Anno sofort als solches zu erkennen.}
The audio file was converted from FLAC to WAV via FFmpeg v4.1.4 \citep{ffmpeg} to meet the aligner's requirements.
This WAV file, the transcription of sentences, and the updated dictionary were
submitted to the aligner that first trained an acoustic model on the data and
then performed the alignment.
In a first step, the resulting .TextGrid file that contains onsets and offsets of words and phonemes was corrected in Praat, again interatively in several passes.
Onsets and offsets of words and their corresponding phonemes on which the automatic alignment performed moderately were corrected.
Some (low volume) sentences that are spoken in continuously noisy settings (pursuit, battle, hurricane, rally, disco) were removed due to poor overall alignment performance.
In a second step, the complete sentences of the orthographic transcription were copied into the annotation (s. table \ref{tab:overview} for an overview of the annotation's content).
In a third step, the name of the speaker was added for every sentence (cf. table \ref{tab:speakers} for the most often occurring speakers).
During every new step results from previous steps were checked again for errors and corrected.

\begin{table*}[tbp]
\caption{Overview of the speech annotation's content for the whole stimulus and it's individual segments used during fMRI scanning. The label ``Sentences''' comprises complete grammatical sentences (i.e. having a subject and a predicate) or questions that are additionally marked with a full stop or question mark.
The  label ``Sentences'' also comprises exlamations or non-speech vocalizations in quick succession (e.g. ``ha, ha, ha'') or in isolation (e.g. ``ha'', at time points when speakers rapidly switch). The label ``Words'' comprises every single word or non-speech vocalization (N=\aPosNonspeechAll) in isolation.}
\label{tab:overview}
\begin{tabular}{llllllllll}
\toprule
\textbf{label} & \textbf{all} & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8}\\
\midrule
sentences  & \aSentencesAll & \aSentencesI & \aSentencesII & \aSentencesIII & \aSentencesIV & \aSentencesV & \aSentencesVI & \aSentencesVII & \aSentencesVIII \tabularnewline
words  & \aWordsAll & \aWordsI & \aWordsII & \aWordsIII & \aWordsIV & \aWordsV & \aWordsVI & \aWordsVII & \aWordsVIII \tabularnewline
phonemes  & \aPhonesAll & \aPhonesI & \aPhonesII & \aPhonesIII & \aPhonesIV & \aPhonesV & \aPhonesVI & \aPhonesVII & \aPhonesVIII \tabularnewline
\bottomrule
\end{tabular}
\end{table*}

\begin{table*}[btp]
\caption{Sentences spoken by the ten most often occurring speakers sorted alphabetically. The narrator only occurrs in the audio-description of the movie. Shown names are mostly identical to the names used in \citep{labs2015portrayed}. Overall 97 persons were identified.}
\label{tab:speakers}
\begin{tabular}{llllllllll}
\toprule
\textbf{name} & \textbf{all} & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8}\\
\midrule
Bubba  & \aBubbaAll & \aBubbaI & \aBubbaII & \aBubbaIII & \aBubbaIV & \aBubbaV & \aBubbaVI & \aBubbaVII & \aBubbaVIII \tabularnewline
Forrest  & \aForrestAll & \aForrestI & \aForrestII & \aForrestIII & \aForrestIV & \aForrestV & \aForrestVI & \aForrestVII & \aForrestVIII \tabularnewline
Forrest (child)  & \aForrestchildAll & \aForrestchildI & \aForrestchildII & \aForrestchildIII & \aForrestchildIV & \aForrestchildV & \aForrestchildVI & \aForrestchildVII & \aForrestchildVIII \tabularnewline
Forrest (v.o.)  & \aForrestvoAll & \aForrestvoI & \aForrestvoII & \aForrestvoIII & \aForrestvoIV & \aForrestvoV & \aForrestvoVI & \aForrestvoVII & \aForrestvoVIII \tabularnewline
Hancock  & \aHancockAll & \aHancockI & \aHancockII & \aHancockIII & \aHancockIV & \aHancockV & \aHancockVI & \aHancockVII & \aHancockVIII \tabularnewline
Jenny  & \aJennyAll & \aJennyI & \aJennyII & \aJennyIII & \aJennyIV & \aJennyV & \aJennyVI & \aJennyVII & \aJennyVIII \tabularnewline
Jenny (child)  & \aJennychildAll & \aJennychildI & \aJennychildII & \aJennychildIII & \aJennychildIV & \aJennychildV & \aJennychildVI & \aJennychildVII & \aJennychildVIII \tabularnewline
Lt. Dan  & \aLtdanAll & \aLtdanI & \aLtdanII & \aLtdanIII & \aLtdanIV & \aLtdanV & \aLtdanVI & \aLtdanVII & \aLtdanVIII \tabularnewline
Mrs. Gump  & \aMrsgumpAll & \aMrsgumpI & \aMrsgumpII & \aMrsgumpIII & \aMrsgumpIV & \aMrsgumpV & \aMrsgumpVI & \aMrsgumpVII & \aMrsgumpVIII \tabularnewline
Narrator  & \aNarratorAll & \aNarratorI & \aNarratorII & \aNarratorIII & \aNarratorIV & \aNarratorV & \aNarratorVI & \aNarratorVII & \aNarratorVIII \tabularnewline
\bottomrule
\end{tabular}
\end{table*}

@@

\todo[inline]{I think this table doesn't talk about 'regressors'. It talks about the
events that were selected/available to build GLM regressors from them. Hence the term
'event' should be used here, and cross-reference to figure \ref{fig:reg} should be made}
\todo[inline]{what does e1-12 stand for? It makes the label column too wide}

@@

pis & substituting indefinite pronoun & \rPisAll & \rPisI & \rPisII & \rPisIII & \rPisIV & \rPisV & \rPisVI & \rPisVII & \rPisVIII \tabularnewline

@@
